\documentclass[12pt]{article}

% bibliography
\usepackage[square,numbers]{natbib}

% mathematics
\usepackage{amssymb}
\usepackage{amsmath}
\newtheorem{theorem}{Theorem}
\newtheorem{proposition}{Proposition}
\newtheorem{remark}{Remark}
\newtheorem{lemma}{Lemma}
\newtheorem{definition}{Definition}

% geometry
\usepackage{hyperref}
\usepackage{setspace}
\usepackage[top=1.2in,bottom=1.2in,left=1.2in,right=1.2in]{geometry}
\onehalfspacing
\hypersetup{
	colorlinks=true,
	urlcolor=blue,
	citecolor=blue,
	linktoc=all,
	linkcolor=red} 

% add your packages if you need



\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% TITLE %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title{\textsc{Your title (limited to 180 characters in length)}} % Title: Limited to 180 characters in length. Use title capitalization.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% AUTHORS %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


\author{Author 1\footnote{Full Address of Author 1. email: email of author 1.}, Author 2\footnote{Full Address of Author 2. email: email of author 2.}
and Author 3\footnote{Full Address of Author 3. email: email of author 3. }\\  % Type all authors' names with initial caps.
}

\date{\empty}
\maketitle


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% ABSTRACT %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{abstract}
Write your abstract which does not exceed 120 words. Write your abstract which does not exceed 120 words. Write your abstract which does not exceed 120 words. Write your abstract which does not exceed 120 words. Write your abstract which does not exceed 120 words. Write your abstract which does not exceed 120 words. 
\end{abstract}

\noindent\textbf{Keywords}: {keyword 1; keyword 2; keyword 3; keyword 4; keyword 5} % Provide up to 5 (non-capitalized) keywords that do not appear in the title. Use semicolons to separate the keywords.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% BODY OF THE PAPER %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Section 1} \label{sec}

The paper body is limited to 600 words, excluding abstract and references, and a maximum of two figures and tables. Use title capitalization for section headings. The paper should be divided into sections, you can include equations, theorems, etc, for example:
\begin{theorem}
Statement. Statement. Statement. Statement. Statement. Statement. Statement. Statement. Statement.
\end{theorem}
References are included at the end of the paper according to the format provided below, and they should be sorted alphabetically. References  are numbered and can be cited as \cite{reference1}.

Write your paper. Write your paper. Write your paper. Write your paper. Write your paper. Write your paper. Write your paper. Write your paper. Write your paper. Write your paper. Write your paper.

\section{Section 2}

Write your paper. Write your paper. Write your paper. Write your paper. Write your paper. Write your paper. Write your paper. Write your paper. Write your paper. Write your paper. Write your paper. 


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% REFERENCES %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{thebibliography}{99}
\bibitem{reference1} Author 1, Author 2  and Author 3 (Year). Title. \textit{Journal}, \textbf{volume}, pp--pp.
\bibitem{reference2} Author 1, Author 2  and Author 3 (Year). \textit{Book title}. Publisher.
\end{thebibliography}

\end{document}
