Two central questions in many probabilistic clustering problems is how many
distinct clusters are present in a particular dataset, and which observations
cluster together.  Bayesian nonparametrics (BNP) addresses this question by
placing a generative process on cluster assignment, making the number of
distinct clusters present amenable to Bayesian inference. However, like all
Bayesian approaches, BNP requires the specification of a prior, and this prior
may favor a greater or fewer number of distinct clusters.  In practice, it is
important to quantitatively establish that the prior is not too informative,
particularly when---as is often the case in BNP---the particular form of the
prior is chosen for mathematical convenience rather than because of a considered
subjective belief.

We derive local sensitivity measures based on Taylor series approximations for a
truncated variational Bayes approximation based on the Kullback-Leibler
divergence.  Using a stick-breaking representation of a Dirichlet process, we
consider perturbations both to the scalar concentration parameter and to the
functional form of the stick-breaking distribution by embedding the
stick-breaking density in the $L_p$ vector spaces of integrable functions.  We
prove that, though the variational optimum is directional differentiable for all
$1 \le p \le \infty$, the derivative provides a uniformly good approximation in
a neighborhood of the original prior only with multiplicative perturbations and
$p=\infty$.

We apply our methods to several real-world datasets, estimating sensitivity to
the BNP prior specification of key posterior quantities, and evaluating the
accuracy of our approximations by comparing to the much more expensive process
of re-fitting the model.  In the design and evaluation of our local sensitivity
measures we pay special attention to our ability to accurately extrapolate to
different priors, rather than treating the sensitivity as a measure of
robustness {\em per se}.  We show the accuracy of our approximation both for
particular parametric and non-parametric perturbations, and demonstrate the
usefulness of the influence function to find maximally influential alternative
priors.
