Recall from \exref{alpha_perturbation} and \corref{our_approximation} derive
sensitivity measures for perturbations that lie within the
$\mathrm{GEM}(\alpha)$ family.  However, there is typically no {\em a priori}
reason to believe that the stick breaking prior lies within this parametric
family.  In this section, we consider perturbing the functional form of the
prior, ultimatlely proving an analogue of \thmref{etat_deriv} for a particular
function space.

To define functional perturbations, fix for the moment a base stick
distribution, $\pbase(\nuk)$ and an alternative stick distribution $\palt(\nuk)$.
Take $\t = \epsilon$ and define
%
\begin{align}\eqlabel{epsilon_pert}
%
\pstick(\nuk \vert \epsilon) ={}&
\frac{\pbase(\nuk)^{1 - \epsilon} \palt(\nuk)^\epsilon}
     {\int_0^1 \pbase(\nuk')^{1 - \epsilon} \palt(\nuk')^\epsilon d\nuk'}.
%
\end{align}
%
We thus have that $\epsilon$ parameterizes a multiplicative path from
$\pbase(\nuk)$ to $\palt(\nuk)$, with $\pstick(\nuk \vert \epsilon = 0) = \pbase(\nuk)$
and $\pstick(\nuk \vert \epsilon=1) = \palt(\nuk)$.  We will take
$\t_0$ to be $\epsilon = 0$, so that we are computing $\etaopt$ with the
prior $\pbase(\nuk)$ and approximating the optimum if we had used $\palt(\nuk)$.
%
Given this definition,
%
\begin{align*}
%
\log \pstick(\nuk \vert \epsilon) ={}&
    \log \palt(\nuk) + \epsilon \log \frac{\palt(\nuk)}{\pbase(\nuk)} + \const.
    & \constdesc{\nuk}
% \\
% \fracat{\partial \log \pstick(\nuk \vert \epsilon) }{\partial \epsilon}{0} ={}&
%     \log \frac{\palt(\nuk)}{\pbase(\nuk)} + \const.
%     & \constdesc{\nuk}
%
\end{align*}
%
Identifying $\t$ with $\epsilon$, we can apply \thmref{etat_deriv} to
$\pstick(\nuk \vert \epsilon)$ as long as $\log (\palt(\nuk) / \pbase(\nuk))$
satisfies \assuitemref{kl_opt_ok}{q_stick_regular}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{ex}
%
We can use \eqref{epsilon_pert} to form an equivalent representatoin of the
parametric perturbation given in \exref{alpha_perturbation}.  For fixed
$\alpha_0$ and $\alpha_1$, take
%
\begin{align*}
%
\pbase(\nuk) :={}& \betadist{\nuk \vert 1, \alpha_0}\\
\palt(\nuk) :={}& \betadist{\nuk \vert 1, \alpha_1}.
%
\end{align*}
%
Then, up to a constant not depending on $\nuk$,
%
\begin{align}\eqlabel{gem_epsilon_pert}
%
\log \pstick(\nuk \vert \epsilon) =
    (\alpha_0 - 1) \log(1 - \nuk) +
    \epsilon  (\alpha_1 - \alpha_0) \log(1 - \nuk).
%
\end{align}
%
Comparing \eqref{gem_alpha_pert} and \eqref{gem_epsilon_pert}, we see by taking
$\alpha = \epsilon \alpha_1$ that the two parameterizations of the prior are
formally equivalent.
%
\end{ex}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

There are many (an infinite number!) of $\palt(\nuk)$ to choose from, so, rather
than fixing $\palt$, it can be useful to think of $\etaopt(\epsilon)$ as {\em
functional} of $\palt$ \citep{gustafson:1996:local}.  Let us fix $\pbase(\nuk)$,
define $\phi(\nuk) := \epsilon \log \left(\palt(\nuk) / \pbase(\nuk)\right)$,
and re-write $\pstick(\nuk \vert \epsilon)$ in the following equivalent form:
%
\begin{align}\eqlabel{phi_perturbation}
%
\pstick(\nuk \vert \phi) ={}&
\frac{\exp\left(\log \pbase(\nuk) + \phi(\nuk)\right)}
     {\int_0^1 \exp\left(\log \pbase(\nuk') + \phi(\nuk')\right) d\nuk'}.
%
\end{align}
%
The advantage of using $\phi$ in \eqref{phi_perturbation} rather than specifying
$\palt$ directly is that $\pstick(\nuk \vert \phi)$ is well-defined for any
Lebesgue-integrable function $\phi(\nuk): (0,1)\mapsto\mathbb{R}$, and a valid
distribution for any function $\phi$ such that the numerator is positive and the
denominator is nonzero.  We can thus safely define perturbations in terms of
general functions without worrying about whether the functions are valid
distributions. We will take $\phiz$ to be the identically zero perturbation so
that $\pstick(\nuk \vert \phiz) = \pbase(\nuk)$.

By analogy with \eqref{kl_shorthand}, we can write
%
\begin{align}\eqlabel{kl_fun_shorthand}
%
\KL{\eta, \phi} := \KL{\q(\zeta \vert \eta) || p(\x \vert \theta, \phi)}
\mathand
\etaopt(\phi) := \argmin_{\zeta \in \etadom} \KL{\eta, \phi}.
%
\end{align}
%
Further, by analogy with \eqref{etalin_def}, if the map $\phi \mapsto
\etaopt(\phi)$ is sufficiently ``nice'', then we might hope to find a linear
functional $d\etaopt(\phi) / d\phi$ such that
%
\begin{align}\eqlabel{etalin_fun_def}
%
\etalin(\phi) := \etaopt + \fracat{d \etaopt(\phi)}{d \phi}{\t=0} \phi
%
\end{align}
%
and $\etaopt(\phi) \approx \etalin(\phi)$ when $\phi$ is ``small'' in some sense
(again, to be specified precisely below).  The compuationally intensive task of
finding $\etaopt(\phi)$ for infinitely many alternative prior distributions can
then be replaced by the analytically and computationally simple $\etalin(\phi)$.
In addition to approximating the effect of particular perturbations, we will see
below that the linearization will allow us to find worst-case perturbations as
well as convenient visual summaries of the effect of prior perturbation.

Our task, then is to find notions of ``nice'' and ''small'' such that
\eqref{etalin_fun_def} holds and $\etaopt(\phi) \approx \etalin(\phi)$. To that
end, we will embed $\phi$ in a complete normed vector space, that is, in a
Banach space.  We can then appeal to the Banach space version of the
implicit function theorem, in perfect analogy with \thmref{etat_deriv}.

Define the norm
%
\begin{align*}
%
\norminf{\phi} :={} \esssup_{\nuk} |\phi(\nuk)|,
%
\end{align*}
%
where $\esssup$ is defined with respect to the Lebesgue measure on $[0,1]$. As
pointed out by \citet{gustafson:1996:local}, $\norminf{\cdot}$ is attactive
because $\pstick(\nuk \vert \phi)$ is a valid distribution whenever $\phi$ lies
within a neighborhood of $\phiz$, as we state in the following
proposition.\footnote{\citet{gustafson:1996:local} also observes that the
converse of \propref{pert_well_defined} is not true---there exist valid
distributions with infinte $\norminf{\phi}$.  We will discuss this and other
issues related to \citet{gustafson:1996:local} below.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{prop}\proplabel{pert_well_defined}
%
% Linf also in \citet[Section 2.1, Example 5]{luenberger:1997:optimization}
Let $\linf$ denote the Banach space of functions on $[0,1]$ with the norm
$\norminf{\cdot}$ \citep[Theorem 5.2.1]{dudley:2018:real}.
There exists a $\norminf{\cdot}$-neighborhood of $\phiz$, denoted $\ball_\phi
\subset \linf$, such that $\pstick(\nuk \vert \phi)$ as defined in
\eqref{phi_perturbation} is a valid probability density for all $\phi \in
\ball_\phi$
%
\begin{proof}
%
The function $\pstick(\nuk \vert \phi)$ is a valid density whenever
$\pstick(\nuk \vert \phi) \ge 0$ with Lebesgue probability one, and $\int
\pstick(\nuk \vert \phi) = 1$. Since $\pbase(\nuk) \ge 0$,
%
\begin{align*}
%
\essinf_{\nuk \in [0,1]} \exp\left(\log \pbase(\nuk) + \phi(\nuk)\right)
    \ge{}&
\essinf_{\nuk \in [0,1]} \pbase(\nuk)
\essinf_{\nuk \in [0,1]} \exp\left(\phi(\nuk)\right)
\ge{}0,
%
\end{align*}
%
so $\pstick(\nuk \vert \phi) \ge 0$ with Lebesgue probability one. Similarly,
since $\int \pbase(\nuk) d\nuk = 1$,
%
\begin{align*}
%
\exp(-\norminf{\phi}) \le{}
\abs{\int_0^1 \exp\left(\log \pbase(\nuk) + \phi(\nuk)\right) d\nuk}
\le{}
\exp(\norminf{\phi}).
%
\end{align*}
%
so that $\int \pstick(\nuk \vert \phi) = 1$ whenever $\norminf{\phi} < \infty$.
%
\end{proof}
%
\end{prop}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Thanks to \propref{pert_well_defined}, we might expect to sensibly write
$\etaopt(\phi) := \argmin_{\eta \in \etadom} \KL{\eta, \phi}$ for the optimal
variational parameters when using the prior $\pstick(\nuk \vert \phi)$.  The
optimum $\etaopt(\phi): \linf \mapsto \mathbb{R}^\etadim$ is thus a functional
of $\phi$.  We might hope to form a Taylor series expansion of $\etaopt(\phi)$.

%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \begin{assu}\assulabel{kl_stick_opt_ok}
% %
% Let the following assumptions hold for some neighborhood $\ball_\eta$
% of $\etaopt(\phiz)$, and for $\eta \in \ball_\eta$:
% %
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \begin{enumerate}
% %
%     \item \itemlabel{kl_diffable}
%     The map $\eta \mapsto \KL{\eta, \phiz}$ is twice
%     continuously differentiable.
% %
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \item\itemlabel{kl_hess}
%     The Hessian matrix $\fracat{\partial^2 \KL{\eta, \phiz}}
%                     {\partial \eta \partial \eta^T}
%                     {\eta}$ is positive definite.%
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \item\itemlabel{q_stick_regular} The variational approximations $\q(\nu
% \vert \etanuk)$ to the stick-breaking posteriors satisfy \assuref{q_regular}
% with $\eta_0 = \etaopt(\phiz)$.
% %
% \end{enumerate}
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% %
% \end{assu}
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{thm}
%
Let \assuref{kl_opt_ok} hold. Then the map $\phi \mapsto \etaopt(\phi)$ is
well-defined and continuously Fr{\'e}chet differentiable in a neighborhood of
$\phiz$ as a map from $\linf$ to $\mathbb{R}^\etadim$. Furthermore, the
derivative in the direction $\phi - \phiz$ is of the formfassuref
%
\begin{align*}
%
\fracat{d \etaopt(\phi)}{d\phi}{\phiz}\left( \phi - \phiz\right) ={}&
    \int \infl(\nu) \phi(\nu) d\nu \mathtxt{where} \\
%
\infl(\nu) :={}&
-\left(\fracat{\partial^2 \KL{\eta, \phiz}}
                {\partial \eta \partial \eta^T}
                {\etaopt}\right)^{-1}
\left(
    \sumkm
    \q(\nu \vert \etaoptnuk)
    \nabla \log \q(\nu \vert \etaoptnuk)
\right).
%
\end{align*}
%

%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{proof}
%
The map $\eta, \phi \mapsto \KL{\eta, \phi}$ is a map from the Banach space
$\mathbb{R}^\etadim \times \linf$ into the Banach space $\mathbb{R}$. Let us
take the L2 norm $\norm{\cdot}_2$ on $\mathbb{R}^{\etadim}$ and $\mathbb{R}$.
Let $\ball$ denote the ball $\ball_\eta \times \{ \phi: \norminf{\phi} <
\delta\}$ for some $\delta > 0$.  Let $\linop$ denote a linear operator from
$\ball$ to $\mathbb{R}^\etadim$, and define the dual norm
%
\begin{align*}
%
\norm{\linop}^* :=
    \sup_{\eta: \norm{\eta}_2 \le 1} \sup_{\phi: \norminf{\phi} \le 1}
     \norm{\linop(\eta, \phi)}_2.
%
\end{align*}

Recall that $\etaopt(\phi)$, if it exists, is defined as a solution to
%
\begin{align*}
%
\fracat{\partial \KL{\eta, \phi}}
                {\partial \eta}
                {\etaopt(\phi)} = 0_\etadim.
%\KLgrad{\etaopt(\phi), \phi} ={}& 0_\etadim.
%
\end{align*}
%
Using this estimating equation, the key to the proof will be the implicit
function theorem for Banach spaces. To use the implicit function theorem, we
need to show that $\KLgrad{\eta, \phi}$ is continuously Fr{\'e}chet
differentiable at $\etaopt(\phiz), \phiz$ and that $\KLhess{\eta, \phi}$ is
positive definite in $\ball$.

Let us define some compact notation for the duration of the proof.
Let
%
\begin{align*}
%
\KLgrad{\eta, \phi} :={}&
    \fracat{\partial \KL{\eta, \phi}}{\partial \eta}{\eta, \phi}
\mathtxt{and}
\KLhess{\eta, \phi} :={}&
    \fracat{\partial^2 \KL{\eta, \phi}}
           {\partial \eta \partial \eta^T}{\eta, \phi}.
%
\end{align*}
%
Additionally, define $\rho(\etanuk, \phi) :={} \expect{\q(\nu \vert
\etanuk)}{\phi(\nu)}$, with
%
\begin{align*}
%
\rho_\eta(\etanuk, \phi) :={}
    % \fracat{\partial \expect{\q(\nu \vert \etanuk)}{\phi(\nu)}}
    %        {\partial \etanuk}{\etanuk} \mathand
\fracat{\partial \rho(\etanuk, \phi)}
       {\partial \etanuk}{\etanuk} \mathand
\rho_{\eta\eta}(\etanuk, \phi) :={}
   \fracat{\partial^2 \rho(\etanuk, \phi)}
          {\partial \etanuk \partial \etanuk^T}{\etanuk}.
%
\end{align*}

Observe that the estimating equation can be written as
%
\begin{align*}
%
% \KL{\eta, \phi} ={}&
%     \KL{\eta, \phiz} + \sumkm \rho(\etanuk, \phi)
%     \Rightarrow \\
%
\KLgrad{\eta, \phi} ={}&
\KLgrad{\eta, \phiz} + \sumkm \rho_\eta(\etanuk, \phi)
= 0_\etadim.
%
\end{align*}

Let us first show that $\KLgrad{\eta, \phi}$ is continuously Fr{\'e}chet
differentiable.  By \assuitemref{kl_opt_ok}{kl_diffable}, $\eta \mapsto
\KLgrad{\eta, \phiz}$ is continuously differentiable and does not depend on
$\phi$, so we need only to show that $\rho_\eta(\etanuk, \phi)$ is continuously
Fr{\'e}chet differentiable. To show this, it will suffice to show that the
partial deritatives of the maps $\eta \mapsto \rho_\eta{\eta, \phi}$ and $\phi
\mapsto \rho_\eta{\eta, \phi}$ exist and are continuous in the dual norm
$\norm{\cdot}^*$ as a function of the location $\eta, \phi$ at which the
derivatives are evaluated.

We first show that $\phi \mapsto \rho_\eta(\etanuk, \phi)$ is Fr{\'e}chet
differentiable.  To do so it


It follows from \eqref{q_sens_is_cov} that the map $\phi \mapsto
\rho_\eta(\etanuk, \phi)$ is a linear functional of $\phi$.  Applying Jensen's
inequality to the $\norm{\cdot}_2$ norm and the triangle inequality twice gives:
%
\begin{align*}
%
%\sup_{\phi: \norminf{\phi} = 1}
\norm{\rho_\eta(\etanuk, \phi)}_2 \le
    4 \expect{\q(\nu \vert \etanuk)}
             {\norm{\nabla \log \q(\nu \vert \etanuk)}_2}
              \norminf{\phi(\nu)}.
%
\end{align*}
%
It follows from ... that $\rho_\eta(\etanuk,
\phi)$ is bounded as a function of $\phi$ for all $\eta \in \ball_\eta$.
Bounded linear functionals are Fr{\'e}chet differentiable (essentially by
definition), so $\phi \mapsto \rho_\eta(\etanuk, \phi)$ is Fr{\'e}chet
differentiable.

It remains to show that the derivative is continuous.

% so the linear operator is bounded as a function of $\phi$, and a continuous
% funciton of $\eta$ by assumption.  Therefore $\phi \mapsto \fracat{\partial
% \expect{\q(\nu \vert \etanuk)} {\phi(\nu)}}{\partial \eta}{\eta}$ is
% continuously Fr{\'e}chet differentiable.
%
% Again differentiating under the integral,
% %
% \begin{align*}
% %
% \fracat{\partial^2 \expect{\q(\nu \vert \etanuk)}
%               {\phi(\nu)}}{\partial \eta \partial \eta^2}{\eta} ={}&
% \expect{\q(\nu \vert \etanuk)}
%        {\left(\nabla \log \q(\nu \vert \etanuk)
%          - \expect{\q(\nu \vert \etanuk)}
%                   {\nabla \log \q(\nu \vert \etanuk)}
%        \right)^2
%        \left( \phi(\nu) - \expect{\q(\nu \vert \etanuk)}{\phi(\nu)} \right)} +
% \\&
% \expect{\q(\nu \vert \etanuk)}{
%        \left(                            \nabla^2 \log \q(\nu \vert \etanuk)
%         - \expect{\q(\nu \vert \etanuk)}{\nabla^2 \log \q(\nu \vert \etanuk)}
%        \right)
%        \left( \phi(\nu) - \expect{\q(\nu \vert \etanuk)}{\phi(\nu)} \right)
%        }.
% %
% \end{align*}
% %
Again this is a linear operator as a function of $\phi$, which is continuous by
assumption.  Consequently the Hessian is invertible in a neighborhood.

It follows by \citet[Proposition 4.14(c)]{zeidler:2013:functional} that
$\KLgrad{\eta, \phi}$ is continuosly Fr{\'e}chet differentiable.  Furthermore,
\citet[Chapter 4 Condition 21b]{zeidler:2013:functional} holds since
$\KLhess{\etaopt(\phiz), \phiz}$ is invertible.  So we satisfy conditions (i),
(ii), and (iii) of \citet[Theorem 4.B(c)]{zeidler:2013:functional}, giving that
the function $\etaopt(\phi)$ exists.  Moreover, since $\KLgrad{\eta, \phi}$ is
continuously Fre{\'e}chet differentiable in a neighborhood of $\etaopt(\phiz),
\phiz$, by \citet[Theorem 4.B(d)]{zeidler:2013:functional}, $\etaopt(\phi)$ is
also continuously Fr{\'e}chet differentiable.

The directional derivative

% Other references: We will use the
% formulation given in \citep[Theorem 3.4.10]{krantz:2012:implicit}, for which
% we must show that... actually that doesn't give differentiability.
% The result follows then from \citet[Proposition 4.8(c)]{zeidler:2013:functional}.
% See also \citet[Corollary 1.4]{averbukh:1967:theory} and \citep[Appendix
% A]{reeds:1976:thesis}).
%
\end{proof}
%
\end{thm}
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
